### 2020-08-26 spinifex_RJpostProofing_20200824/gifs/jetcluster_radialtour_pc4.rmd

---
title: "Supplemental material -- figure 5 animation: jet cluster, radial tour of PC4"
output: html_document
---

```{r, echo=FALSE, fig.cap="Animation of figure 5. Snapshots of a radial manual tour of PC4 within the jet cluster, with color indicating experiment type: ATLAS7new (green) and ATLAS7old (orange). When PC4 is removed from the projection (frame 10) there is little difference between the groups, suggesting that PC4 is important for distinguishing the experiments."}
knitr::include_graphics("../gifs/jetcluster_radialtour_pc4.gif")
```